---
title: "PML Final Project"
output: html_document
---

        
        Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).
Data

The training data for this project are available here:
        
        https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
        
        https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

Building off of the previous instructions, the analysis commenced by loading the required packages:
```{r}
#library(R2HTML)
library(caret)
#library(doParallel)
library(randomForest)
```

```{r}
# Download data.
url_raw_training <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
file_raw_training <- "pml-training.csv"
download.file(url=url_raw_training, destfile=file_raw_training, method="curl")
url_raw_testing <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
file_raw_testing <- "pml-testing.csv"
download.file(url=url_raw_testing, destfile=file_raw_testing, method="curl")
```
To allow for cross validation of the dataset the data are kept separate as training and testing sets. Utilizing cross validation allows for preservation of the testing dataset to ensure independence and accuracy in the final results.
```{r}
training<-read.csv(file_raw_training, na.strings=c("", "NA"))
testing<-read.csv(file_raw_testing)
```

```{r}
training$X <- NULL
train <- training[,colSums(is.na(training)) < 19216]
```

A Random Forest algorithm was chosen for the prediction function. This was due to the increased accuracy afforded over other option when using an RF model. Using the train function from the caret package to fit the model with a principal components pre Processing algorithm. The downside to a random forest was the computational time required to produce the model. This could be reduced using multi-core processor.
```{r}
rfModel <- train(classe ~ .,  method="rf", preProcess="pca", trControl=trainControl(method = "cv", number = 10), data=train)
rfModel
```
In-Sample Error

The prediction algorithm was used against the training dataset to determine the accuracy of the model.
```{r}
pT <- predict(rfModel, train)
table(pT, train$classe)
```
The prediction algorithm's in-sample error was 100% as reported above. Given that the training set produced the algorithm, the high accuracy is predictable, though still higher than expected. The out of sample error is the critical piece and that was run next.

Testing against 20 cases in the test set, the resulting output was used to determine the out of sample error.
```{r}
predT <- as.character(predict(rfModel, testing))
predT
```
When submitting these 20 results to the website, the algorithm correctly predicted all 20 test cases, yielding an out of sample error of 100%. As a first pass, this result is highly encouraging, though further testing on a new data, larger, dataset would potentially yield different, less accurate, results.

The random forest algorithm was chosen as the most accurate modelling type based on the information presented in class. While other model types could be run to determine their accuracy, the computational time precluded the use of multiple types - especially given the results of the prediction from the RF model. 